\bf Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-03-16 doc/paper.Rmd
couple of edits

\hh{Here's a bit of a jump: you're already talking about features, but we haven't yet properly introduced signatures of bullets yet.}
Bullet matching methods depend on the attributes  used for making comparisons. For single feature based matching \citet{chu2013} used an automatic method for counting consecutive matching striae (CMS). The authors report an error rate of 52% of the known same source lands comparisons as misidentified (false negative) and zero false positives for known different source lands. \citet{ma2004} and \citet{vorburger2011} discuss CCF (cross-correlation function) and its discriminating power and applicability for same-source analyses of bullets, but do not provide any error rates in their discussion. \citet{aoas} use multiple features like CCF, CMS, D (distance measure) etc in a random forest based method and compare every land against every other land of  digitised versions of Hamby 252 and Hamby 44 \citep{hamby} published on the NIST Ballistics Database \citep{nist}. The authors report an out-of-bag overall error rate of 0.46%, comprised of a false positive error rate of 30.05% and a false negative rate of 0.026%.

The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks. \citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same tool. The data for their study was obtained from 50 sequentially manufactured screwdriver tips. \citet{chumbley} reported error rates for marking made by the tips at differnt angles. For a 30 degree they report false negative error rate to be 0.023 and  false positive error rate be 0.09. For other angles the error rates remain the same for false negatives and 0.01 for false positives. The improvement proposed by \citet{hadler} on the other hand, did not consider effect of angles and gave 0.06 false  negative error rate and 0 false poisitive error rate for markings made at the same angle.

@@

The land engraved markings across the cross section of the land are termed as bullet profiles \citep{aoas,ma2004}. These are often used for making land to land comparisons. Bullet signatures on the other hand \citep{chu2013,aoas} refer to a processed version of the raw land engraved markings or profiles. The generation of bullet signatures involves first extraction of a bullet profile by taking the cross-sectional of the surface at a given height. Loess fits are then used to model the structure. The residuals of these fits are called signatures. They are considered to be noise free and a good reflection of the key attributes of the raw marking, and the unique features of a bullet. A more detailed version of the extraction technique of signatures is discussed by \citet{aoas}, where comprehensive details about the height at which profile is to be selected, removing curvature, smoothing, identifying groove locations are explained.

The majority of Bullet profiles and signatures of the Hamby study \citep{hamby} used in this paper are extracted by procedures mentioned by  \citet{aoas}. The scans for the Hamby study come from two sources namely, NIST scans \citep{nist} and scans produced at Iowa State University. Both scans were generated using confocal microscopes but at different resolutions. The NIST database contains scans of the bullet sets Hamby 44 and 252 with the resolution 1.5625 $\mu m$ per pixel. The Iowa State scans are available for only the Hamby 44 bullet set and at a resoultion of 0.65 $\mu m$ per pixel. The data used by \citet{chumbley, hadler} on the other hand, was produced by a profilometer at a resolution of about 0.73 $\mu m$ per pixel. The different resolutions indicate that the length of the digitized markings would be significantly different for each setting.

@@

Markings made by Screw driver tips \citep{chumbley} are longer and pronounced and about 7 mm in length \citep{manytoolmarks1}, while the groove to groove length of the bullet lands from the Hamby study \citep{hamby} are about 2mm in length. The Ruger pistol barrels used in the  Hamby study have a 9mm caliber. This means that bullets used in these pistols have land markings of almost $(1/4)^{th}$ the size of screwdriver toolmarks used by \citet{chumbley, hadler}, therefore raising a problem of distinctiveness. One way to investigate this is to evaluate bullet striations at different resolutions. This gives us an incentive to use both the NIST and Iowa State scans and investigate the effect of length of the digitized mark on the working of the algorithm. Apart from this, the algorithm developed by \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature, whereas striations on bullets are made on their curved surfaces. Therefore, using methods proposed for toolmarks need adaptation in order to give tangible results for bullets. Another challenge that arises from these topographical differences, is the understanding of that effect of different levels of smoothing. Especially investigating its effect on the algorithm performance when signatures are being compared and when raw markings or profiles are being compared.

>>>>

Bullet matching methods are usually based on these associated signatures. \citet{chu2013} use an automatic method for counting consecutive matching striae (CMS). The authors report an error rate of 52% of the known same source lands comparisons as misidentified (false negative) and zero false positives for known different source lands. \citet{ma2004} and \citet{vorburger2011} discuss CCF (cross-correlation function) and its discriminating power and applicability for same-source analyses of bullets, but do not provide any error rates in their discussion. \citet{aoas} use multiple features like CCF, CMS, D (distance measure) etc in a random forest based method and compare every land against every other land of  digitised versions of Hamby 252 and Hamby 44 \citep{hamby} published on the NIST Ballistics Database \citep{nist}. The authors report an out-of-bag overall error rate of 0.46%, comprised of a false positive error rate of 30.05% and a false negative rate of 0.026%.

The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks. \citet{chumbley} compare two toolmarks for same-source. The data for this study was obtained from 50 sequentially manufactured screwdriver tips. \citet{chumbley} report error rates for markings made by the tips at different angles. For markings made under a 30 degree the authors  report an average false negative error rate of 0.023  and an average false positive error rate of 0.09. For other angles the error rates for false negatives stay the same while the rate of false positives decreases to 0.01. \hh{are the angles steeper?} 
The paper by \citet{hadler} is based on the same data but the authors focus on  markings made under the same angle. The error rates associated with the deterministic version of the score are  0.06 for false  negatives and a false poisitive error rate of 0.

@@

Comparisons of striae from bullets are usually based on comparisons of striae in land engraved areas, which are extracted in form of cross sections, called *profiles* \citep{aoas,ma2004}. From profiles bullet *signatures* \citep{chu2013,aoas} are extracted as residuals of a loess fit (or Gaussian filter).
 They are considered to be noise free and a good reflection of the key attributes of the raw marking, and the unique features of a bullet. \hh{Do you have a reference for the previous sentenced?} A  detailed discussion of the extraction technique for signatures is given in \citet{aoas}.
 <!--, where comprehensive details about the height at which profile is to be selected, removing curvature, smoothing, identifying groove locations are explained.-->

\hh{don't split the discussion on the size. between the next paragraph and }
<!--The majority of Bullet profiles and signatures of the Hamby study \citep{hamby} used in this paper are extracted by procedures mentioned by  \citet{aoas}.--> The scans for the Hamby study come from two sources namely, NIST scans \citep{nist} and scans produced at Iowa State University \hh{Iowa State is too general. it's the High-resolution microscopy lab and we are CSAFE}. Both scans were generated using confocal light microscopes at a 20x magnification but the microscopes collected data at different resolutions. The NIST database contains scans of the bullet sets Hamby 44 and 252 at a resolution of 1.5625 $\mu m$ per pixel. The Iowa State scans are available for only the Hamby 44 bullet set and at a resoultion of 0.645 $\mu m$ per pixel. The data used by \citet{chumbley, hadler} on the other hand, was produced by a profilometer at a resolution of about 0.73 $\mu m$ per pixel. The different resolutions indicate that the length of the digitized markings would be significantly different for each setting.

@@

Markings made by screw driver tips \citep{chumbley} are longer and pronounced and about 7 mm in length \citep{manytoolmarks1}, while the groove to groove length of the bullet lands from the Hamby study \citep{hamby} are about 2mm in length. The Ruger pistol barrels used in the  Hamby study have a 9mm caliber. This means that bullets used in these pistols have land markings of only about a quarter of the size of screwdriver toolmarks used by \citet{chumbley, hadler}, therefore raising a problem of distinctiveness. One way to investigate this is to evaluate bullet striations at different resolutions. This gives us an incentive to use both the NIST and Iowa State scans and investigate the effect of length of the digitized mark on the working of the algorithm. Apart from this, the algorithm developed by \citet{hadler} has only been tested for flatter and wider surfaces which have negligible curvature, whereas striations on bullets are made on their curved surfaces. Therefore, using methods proposed for toolmarks need adaptation in order to give tangible results for bullets. Another challenge that arises from these topographical differences, is the understanding of that effect of different levels of smoothing. Especially investigating its effect on the algorithm performance when signatures are being compared and when raw markings or profiles are being compared.

### 2018-03-16 doc/paper.tex
couple of edits

Various toolmarks have been studied in the literature:
\citet{manytoolmarks1} and \citet{chumbley} have been analyzing
screwdriver marks digitized using a profilometer; \citet{manytoolmarks2}
have investigated 3D marks from screwdriver, tongue and groove pliers
captured using a confocal microscope; \citet{afte-chumbley} have been
investigated digitized marks from slip-joint pliers generated by a
surface profilometer.

\hh{We need an additional sentence here ... } \citet{manytoolmarks2}
define a relative distance metric and use it as similarity measure
between two toolmarks. \citet{manytoolmarks1} extract many small
segments in the markings of two toolmarks and compare similiarity using
a maximum pearson correlation coefficient. The Chumbley scoring method,
first introduced by \citet{chumbley}, uses a similar but more extensive
framework based on a Mann-Whitney U test of the resulting correlation
coefficients. This approach was later improved by Hadler and Morris
\citep{hadler} where the authors introduce a deterministic approach and
address the issue of lack of independence between segements of striae.
In this paper, we are investigating the applicability of the Chumbley
scoring method by \citet{hadler} to assess striation marks on bullet
lands for same-source identification.

@@

\hh{Here's a bit of a jump: you're already talking about features, but we haven't yet properly introduced signatures of bullets yet.}
Bullet matching methods depend on the attributes used for making
comparisons. For single feature based matching \citet{chu2013} used an
automatic method for counting consecutive matching striae (CMS). The
authors report an error rate of 52\% of the known same source lands
comparisons as misidentified (false negative) and zero false positives
for known different source lands. \citet{ma2004} and
\citet{vorburger2011} discuss CCF (cross-correlation function) and its
discriminating power and applicability for same-source analyses of
bullets, but do not provide any error rates in their discussion.
\citet{aoas} use multiple features like CCF, CMS, D (distance measure)
etc in a random forest based method and compare every land against every
other land of digitised versions of Hamby 252 and Hamby 44 \citet{hamby}
published on the NIST Ballistics Database \citep{nist}. The authors
report an out-of-bag overall error rate of 0.46\%, comprised of a false
positive error rate of 30.05\% and a false negative rate of 0.026\%.

The Chumbley score provides us with another approach in the same-source
assessment of bullet striation marks.\citet{chumbley}, in their paper,
compare two toolmarks with the intention of determining if it comes from
the same tool. The data for their study was obtained from 50
sequentially manufactured screwdriver tips.\citet{chumbley} reported
error rates for marking made by the tips at differnt angles. For a 30
degree they report false negative error rate to be 0.023 and false
positive error rate be 0.09. For other angles the error rates remain the
same for false negatives and 0.01 for false positives. The improvement
proposed by \citet{hadler} on the other hand, did not consider effect of
angles and gave 0.06 false negative error rate and 0 false poisitive
error rate for markings made at the same angle.

@@

The land engraved markings across the cross section of the land are
termed as bullet profiles \citep{aoas,ma2004}. These are often used for
making land to land comparisons. Bullet signatures on the other hand
\citep{chu2013,aoas} refer to a processed version of the raw land
engraved markings or profiles. The generation of bullet signatures
involves first extraction of a bullet profile by taking the
cross-sectional of the surface at a given height. Loess fits are then
used to model the structure. The residuals of these fits are called
signatures. They are considered to be noise free and a good reflection
of the key attributes of the raw marking, and the unique features of a
bullet. A more detailed version of the extraction technique of
signatures is discussed by \citet{aoas}, where comprehensive details
about the height at which profile is to be selected, removing curvature,
smoothing, identifying groove locations are explained.

The majority of Bullet profiles and signatures of the Hamby study
\citep{hamby} used in this paper are extracted by procedures mentioned
by \citet{aoas}. The scans for the Hamby study come from two sources
namely, NIST scans \citep{nist} and scans produced at Iowa State
University. Both scans were generated using confocal microscopes but at
different resolutions. The NIST database contains scans of the bullet
sets Hamby 44 and 252 with the resolution 1.5625 \(\mu m\) per pixel.
The Iowa State scans are available for only the Hamby 44 bullet set and
at a resoultion of 0.65 \(\mu m\) per pixel. The data used by
\citet{chumbley, hadler} on the other hand, was produced by a
profilometer at a resolution of about 0.73 \(\mu m\) per pixel. The
different resolutions indicate that the length of the digitized markings
would be significantly different for each setting.

@@

Markings made by Screw driver tips \citep{chumbley} are longer and
pronounced and about 7 mm in length \citep{manytoolmarks1}, while the
groove to groove length of the bullet lands from the Hamby study
\citep{hamby} are about 2mm in length. The Ruger pistol barrels used in
the Hamby study have a 9mm caliber. This means that bullets used in
these pistols have land markings of almost \((1/4)^{th}\) the size of
screwdriver toolmarks used by \citet{chumbley, hadler}, therefore
raising a problem of distinctiveness. One way to investigate this is to
evaluate bullet striations at different resolutions. This gives us an
incentive to use both the NIST and Iowa State scans and investigate the
effect of length of the digitized mark on the working of the algorithm.
Apart from this, the algorithm developed by \citet{hadler} has only been
tested for flatter and wider surfaces which have negligible curvature,
whereas striations on bullets are made on their curved surfaces.
Therefore, using methods proposed for toolmarks need adaptation in order
to give tangible results for bullets. Another challenge that arises from
these topographical differences, is the understanding of that effect of
different levels of smoothing. Especially investigating its effect on
the algorithm performance when signatures are being compared and when
raw markings or profiles are being compared.

>>>>

\hh{In the next page and a half it is easy to lose the red line. It might help to include a table with an overview.
The table should include the reference to the paper, the data used, the statistical method and the associated error rates.}
Various toolmarks have been studied in the literature:
\citet{manytoolmarks1} and \citet{chumbley} have been analyzing
screwdriver marks digitized using a profilometer; \citet{manytoolmarks2}
have investigated 3D marks from screwdriver, tongue and groove pliers
captured using a confocal microscope; \citet{afte-chumbley} have been
investigated digitized marks from slip-joint pliers generated by a
surface profilometer.

\hh{We need an additional sentence here to get from the data to the statistical methods ... }
\citet{manytoolmarks2} define a relative distance metric and use it as
similarity measure between two toolmarks. \citet{manytoolmarks1} extract
many small segments in the markings of two toolmarks and compare
similiarity using a maximum pearson correlation coefficient. The
Chumbley scoring method, first introduced by \citet{chumbley}, uses a
similar but more extensive framework based on a Mann-Whitney U test of
the resulting correlation coefficients. This approach was later improved
by Hadler and Morris \citep{hadler} where the authors introduce a
deterministic approach and address the issue of lack of independence
between segements of striae. In this paper, we are investigating the
applicability of the Chumbley scoring method by \citet{hadler} to assess
striation marks on bullet lands for same-source identification.

@@

Bullet matching methods are usually based on these associated
signatures. \citet{chu2013} use an automatic method for counting
consecutive matching striae (CMS). The authors report an error rate of
52\% of the known same source lands comparisons as misidentified (false
negative) and zero false positives for known different source lands.
\citet{ma2004} and \citet{vorburger2011} discuss CCF (cross-correlation
function) and its discriminating power and applicability for same-source
analyses of bullets, but do not provide any error rates in their
discussion. \citet{aoas} use multiple features like CCF, CMS, D
(distance measure) etc in a random forest based method and compare every
land against every other land of digitised versions of Hamby 252 and
Hamby 44 \citep{hamby} published on the NIST Ballistics Database
\citep{nist}. The authors report an out-of-bag overall error rate of
0.46\%, comprised of a false positive error rate of 30.05\% and a false
negative rate of 0.026\%.

The Chumbley score provides us with another approach in the same-source
assessment of bullet striation marks. \citet{chumbley} compare two
toolmarks for same-source. The data for this study was obtained from 50
sequentially manufactured screwdriver tips. \citet{chumbley} report
error rates for markings made by the tips at different angles. For
markings made under a 30 degree the authors report an average false
negative error rate of 0.023 and an average false positive error rate of
0.09. For other angles the error rates for false negatives stay the same
while the rate of false positives decreases to 0.01.
\hh{are the angles steeper?} The paper by \citet{hadler} is based on the
same data but the authors focus on markings made under the same angle.
The error rates associated with the deterministic version of the score
are 0.06 for false negatives and a false poisitive error rate of 0.

@@

Comparisons of striae from bullets are usually based on comparisons of
striae in land engraved areas, which are extracted in form of cross
sections, called \emph{profiles} \citep{aoas,ma2004}. From profiles
bullet \emph{signatures} \citep{chu2013,aoas} are extracted as residuals
of a loess fit (or Gaussian filter). They are considered to be noise
free and a good reflection of the key attributes of the raw marking, and
the unique features of a bullet.
\hh{Do you have a reference for the previous sentenced?} A detailed
discussion of the extraction technique for signatures is given in
\citet{aoas}.

\hh{don't split the discussion on the size. between the next paragraph and }
The scans for the Hamby study come from two sources namely, NIST scans
\citep{nist} and scans produced at Iowa State University
\hh{Iowa State is too general. it's the High-resolution microscopy lab and we are CSAFE}.
Both scans were generated using confocal light microscopes at a 20x
magnification but the microscopes collected data at different
resolutions. The NIST database contains scans of the bullet sets Hamby
44 and 252 at a resolution of 1.5625 \(\mu m\) per pixel. The Iowa State
scans are available for only the Hamby 44 bullet set and at a resoultion
of 0.645 \(\mu m\) per pixel. The data used by \citet{chumbley, hadler}
on the other hand, was produced by a profilometer at a resolution of
about 0.73 \(\mu m\) per pixel. The different resolutions indicate that
the length of the digitized markings would be significantly different
for each setting.

@@

Markings made by screw driver tips \citep{chumbley} are longer and
pronounced and about 7 mm in length \citep{manytoolmarks1}, while the
groove to groove length of the bullet lands from the Hamby study
\citep{hamby} are about 2mm in length. The Ruger pistol barrels used in
the Hamby study have a 9mm caliber. This means that bullets used in
these pistols have land markings of only about a quarter of the size of
screwdriver toolmarks used by \citet{chumbley, hadler}, therefore
raising a problem of distinctiveness. One way to investigate this is to
evaluate bullet striations at different resolutions. This gives us an
incentive to use both the NIST and Iowa State scans and investigate the
effect of length of the digitized mark on the working of the algorithm.
Apart from this, the algorithm developed by \citet{hadler} has only been
tested for flatter and wider surfaces which have negligible curvature,
whereas striations on bullets are made on their curved surfaces.
Therefore, using methods proposed for toolmarks need adaptation in order
to give tangible results for bullets. Another challenge that arises from
these topographical differences, is the understanding of that effect of
different levels of smoothing. Especially investigating its effect on
the algorithm performance when signatures are being compared and when
raw markings or profiles are being compared.
